---
title: "Gov 50 Final Project"
author: "Tay'vian Laing"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  markdown: 
    wrap: sentence
---

## Project thoughts

I am interested in exploring data related to the amount of black lawyers through a certain time frame.
Also, I am interested in exploring data related to the amount of black corporate lawyers in top notch law firms or how many black corporate lawyers excel to become senior partner.
Ultimately, I am interested in exploring data around black lawyers.

## Project Research Question & Hypothesis

Does having a better ranked legal representation actually correlate with the success of litigation cases?
In this study, I plan to compare the success of litigators who have better ranked legal representation to those who do not have as good representation.
I hypothesize that better ranked legal representation does indeed correlate with the success of litigation because of the resources/prestige that comes with being represented by the best.
I expect that it's easier for a judge or whomever to side with a litigator from a well represented firm simply because of the renown and prestige of the firm, so the people already know what to expect from these well represented people.
The explanatory variables in this data set are the different types of ranks that lawyers are given in the UK.
These ranks consist of Silk, Band, Junior, and much more.
These ranks will be used for analysis to see how different outcomes can be in our study.
In this study, our outcome variable will be the outcomes in tax litigation.
This will be measured by seeing the results in which higher ranking lawyers successfully litigate compared to lower ranked lawyers.
The observed pattern that will prove my hypothesis would be the results in which better ranked lawyers have successfully litigated.
On the other hand, if more lower ranking lawyers successfully litigate more than higher ranking lawyers, then that would disprove my hypothesis.
In analyzing these patterns, the answer to the research question will be answered.

## Data Visualization

```{r}
library(tidyverse)
library(ggplot2)
library(readr)

library(readr)
union_with_ranks <- read_csv("union-with-ranks.csv")

appellants <- union_with_ranks |>
  ggplot(mapping = aes(x = AppType)) + geom_bar(fill = "indianred1") + labs(x = "Appellant Type", y = "Appellants")
appellants

respondents <- union_with_ranks |>
  ggplot(mapping = aes(x = RespType)) + geom_bar(aes(fill = "indianred1")) + labs(x = "Respondent Type", y = "Respondents")
respondents

```

The graphs above represent the two types of lawyers in this study, the appellants and the respondents.
As shown above, both the respondents and the appellants are mostly backed up by the government in comparison to the other available types.
Another key observation that I noticed is that there are less individual respondent types in comparison to the amount of appellant individual types.

## Analysis

```{r}
apptype <- union_with_ranks |>
  group_by(Outcome, AppType) |>
  pivot_wider(names_from = app.rank1,
               values_from = Outcome) |>
  knitr::kable()
```

To answer my research question, the better represented a litigator is doesn't necessarily mean that there is a correlation between their success in court.
I previously hypothesized that there would be a positive correlation between the rate of success in court and how good one litigator is represented.
However, this shows that my hypothesis was wrong and there is no correlation to be found.
